Understanding the Risk of Casualties in US School Shootings

### 2024-12-04 paper/paper.qmd
update graphs

# Plot the total number of casualties by shooting type
ggplot(casualties_by_type, aes(x = reorder(shooting_type, total_casualties), y = total_casualties, fill = shooting_type)) +
  geom_bar(stat = "identity", color = "white") +
  scale_fill_brewer(palette = "Set3") +
  labs(
    title = "Distribution of Casualty By Shooting Type",
    x = "Shooting Type",
    y = "Total Number of Casualties",
    fill = "Shooting Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),  # Rotate x-axis labels for readability
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.position = "none"  # Hide legend as labels are self-explanatory
  )
```

@@

```{r}
#| label: fig-casualtydist
#| echo: false
#| fig-cap: "Distribution of Casualties in US School Shooting Incidents from 1999 to 2024"
# Causaulity Distribution
ggplot(data, aes(x = casualties)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "white") + 
  labs(title = "Distribution of Casualties", x = "Number of Casualties", y = "Frequency") +
  theme_minimal() +
    theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
    )

@@

# Adjusted Plot
ggplot(map_data, aes(long, lat, group = group, fill = total_casualties)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(
    low = "#E6F7FF",
    high = "#0072B2",
    name = "Total Casualties",
    breaks = c(0, 50, 100, 150, 200)  # Adjust breaks based on your data
  ) +
  labs(
    title = "US School Shooting Casualties Visualization",
    x = NULL,
    y = NULL
  ) +
  coord_fixed(1.3) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14, margin = margin(t = 10, b = 10)),
    plot.subtitle = element_text(hjust = 0.5, size = 12, margin = margin(t = 0, b = 20)),
    legend.position = "right",
    legend.title = element_text(size = 10),  # Reduce the size of "Total Casualties"
    legend.text = element_text(size = 8),   # Reduce size of legend values
    legend.key.height = unit(0.5, "cm"),    # Reduce the height of the legend color bar
    legend.key.width = unit(0.25, "cm"),    # Reduce the width of the legend color bar
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)  # Increase margins for spacing
  ) +
  guides(
    fill = guide_colorbar(
      barwidth = 0.5,  # Adjust the width of the legend bar
      barheight = 8    # Adjust the height of the legend bar
    )
  )

@@

# Plot Pie Chart with Percentages for Top 4
ggplot(shooting_data, aes(x = "", y = proportion, fill = shooting_type)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar(theta = "y") +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(
    title = "Distribution of Shooting Types",
    fill = "Shooting Type"
  ) +
  geom_text(
    aes(label = label), 
    position = position_stack(vjust = 0.5), 
    size = 4, 
    color = "black"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.title = element_text(size = 10),
    legend.position = "right"
  )

@@

# Plot the distribution of latitude
ggplot(data, aes(x = lat)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "white") +
  theme_minimal() +
  labs(
    title = "Distribution of Latitude of Schools in Shooting Incidents",
    x = "Latitude",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  )

>>>>

# Plot the total number of casualties by shooting type
ggplot(casualties_by_type, aes(x = reorder(shooting_type, total_casualties), y = total_casualties, fill = shooting_type)) +
  geom_bar(stat = "identity", color = "white") +
  scale_fill_brewer(palette = "Set3") +
  labs(
    x = "Shooting Type",
    y = "Total Number of Casualties",
    fill = "Shooting Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),  # Rotate x-axis labels for readability
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.position = "none"  # Hide legend as labels are self-explanatory
  )
```

@@

```{r}
#| label: fig-casualtydist
#| echo: false
#| fig-cap: "Distribution of Casualties in US School Shooting Incidents from 1999 to 2024"
# Causaulity Distribution
ggplot(data, aes(x = casualties)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "white") + 
  labs(x = "Number of Casualties", y = "Frequency") +
  theme_minimal() +
    theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
    )

@@

# Adjusted Plot
ggplot(map_data, aes(long, lat, group = group, fill = total_casualties)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(
    low = "#E6F7FF",
    high = "#0072B2",
    name = "Total Casualties",
    breaks = c(0, 50, 100, 150, 200)  # Adjust breaks based on your data
  ) +
  labs(
    x = NULL,
    y = NULL
  ) +
  coord_fixed(1.3) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14, margin = margin(t = 10, b = 10)),
    plot.subtitle = element_text(hjust = 0.5, size = 12, margin = margin(t = 0, b = 20)),
    legend.position = "right",
    legend.title = element_text(size = 10),  # Reduce the size of "Total Casualties"
    legend.text = element_text(size = 8),   # Reduce size of legend values
    legend.key.height = unit(0.5, "cm"),    # Reduce the height of the legend color bar
    legend.key.width = unit(0.25, "cm"),    # Reduce the width of the legend color bar
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)  # Increase margins for spacing
  ) +
  guides(
    fill = guide_colorbar(
      barwidth = 0.5,  # Adjust the width of the legend bar
      barheight = 8    # Adjust the height of the legend bar
    )
  )

@@

# Plot Pie Chart with Percentages for Top 4
ggplot(shooting_data, aes(x = "", y = proportion, fill = shooting_type)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar(theta = "y") +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(
    fill = "Shooting Type"
  ) +
  geom_text(
    aes(label = label), 
    position = position_stack(vjust = 0.5), 
    size = 4, 
    color = "black"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.title = element_text(size = 10),
    legend.position = "right"
  )

@@

# Plot the distribution of latitude
ggplot(data, aes(x = lat)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "white") +
  theme_minimal() +
  labs(
    x = "Latitude",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  )
